---
title: "a2_HunterRaymond_t2"
author: "Ray Hunter"
date: "2023-02-12"
output:   
  html_document:
    code_folding: hide
---

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE,  results=FALSE)
library(purrr)
library(tidyverse)
library(janitor)
library(Metrics)
library(cowplot)
library(here)
library(kableExtra)
```

## Task 2

#### Introduction:

The purpose this task is to answer the question: Do do this, we use a non linear least squares to estimate parameters of a lnegth to wight model for lizard populations in New Mexico. 

All data was collect...



**Dats Source:** Lightfoot, D. and W.G. Whitford. 2020. Lizard pitfall trap data from 11 NPP study locations at the Jornada Basin LTER site, 1989-2006 ver 37. Environmental Data Initiative. https://doi.org/10.6073/pasta/4a6e258fb49c31e222ecbbcfd128967f


### 1: 



We will need to Fit a snout length to weight model of the following form to all lizards in your dataframe.
Since we have a variable from the future we can use the `lead()` function. Make sure to drop the `NA` created by the lead function.

```{r}
mdl_data<-  read_csv(here("data", "lizards.csv")) %>% 
  clean_names() %>% 
  mutate(log_weight = log(weight), log_length = log(sv_length)) 
  
 
```


#### creating a function

```{r }
func1<-function(a,b,length){
 out= a*(length^b)
return(out)
}
```



#### initial guess 

$$  W =a(SVL)^b $$

$$ log(W)= log(a) + b*log(SVL)$$
Because the data has an exponential nature, we need to use a log transformation to creat a linear model using OSL. This then gives us the coefficient values of the model that can be applied to the orignial equation.

```{r}
guess_mod <- lm(log_weight ~log_length, data = mdl_data) 
summary(guess_mod)
```






W = a(SVL)^b

$$  W =a(SVL)^b $$

$$ W = e^{-8.25} (SVL)^{2.48}$$



#### NLS model and figure
```{r}
nls_mdl=nls(weight ~func1(a,b, sv_length),
                  data=mdl_data,
                  start=list(a=exp(-8.25), b=2.479417),
                  trace=TRUE)
```


```{r}

mdl_predict<-mdl_data %>% 
  mutate(predict=predict(nls_mdl,newdata=.))

ggplot1 <- ggplot(data=mdl_predict)+
  geom_point(aes(x=sv_length,y=weight,colour = sex)) + 
    geom_smooth(aes(x=sv_length,y=predict), color='black') +
   facet_wrap(~sex)
  

```







#### NLS model output
```{r}
broom::tidy(nls_mdl) %>% 
  kable(caption = "Original NLS") %>% 
  kable_classic()
```





#### second NLS


```{r}
mdl2_data <- mdl_data %>% 
  filter(spp =="CNTI", sex=="M")
```


#### NLS model and figure
```{r}
nls_mdl2=nls(weight ~func1(a,b, sv_length),
                  data=mdl2_data,
                  start=list(a=exp(-9.0240), b=2.6981),
                  trace=TRUE)
```

```{r}
guess_mod2 <- lm(log_weight ~log_length, data = mdl2_data) 
summary(guess_mod2)
```

```{r}

mdl2_predict<-mdl2_data %>% 
  mutate(predict=predict(nls_mdl2,newdata=.)) 

ggplot2 <- ggplot(data=mdl2_predict)+ xlim(0,100) +ylim(0,40) +
  geom_point(aes(x=sv_length,y=weight)) + 
    geom_smooth(aes(x=sv_length,y=predict), color='blue') 
  
ggplot1
ggplot2
```
































































