---
title: "a2_HunterRaymond_T1"
author: "Ray Hunter"
date: "2023-02-11"
output:   
  html_document:
    code_folding: hide
---

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE,  results=FALSE)
library(tidyverse)
library(here)
library(tidymodels)
library(palmerpenguins)
library(GGally)
library(jtools)
library(AICcmodavg)
library(broom) 
library(knitr)
library(patchwork)
library(kableExtra)
```

## Task 1

#### Introduction:

The purpose this task is to answer the question: Can palmetto plant traits (height, canopy length, canopy width, and leaf count) predict the specific species as *Serenoa repens* or *Sabal etonia*? This was answered by creating two different binary logistic regression models testing the variables' ability to predict the species. Data was collected in south-central Florida from 1981 to 2017 at five year intervals (Abrhamson, 2019). Plant height, caopy length, and canopy width were measure in centimeters (cm) and leaf count was measured by individuals observed.

**Data source:** Abrahamson, W.G. 2019. Survival, growth and biomass estimates of two dominant palmetto species of south-central Florida from 1981 - 2017, ongoing at 5-year intervals ver 1. Environmental Data Initiative.

b.  A section containing 2 - 3 finalized (customized, suitable for a publication) data visualizations (with figure captions) in which you explore differences in height, canopy length, canopy width, and green leaves for the two species. If you prefer, combine the figures into a compound figure using {patchwork} or {cowplot}. Below your data visualizations, add a sentence or two with a takeaway from the plots, e.g., based on these plots, which predictor variables are more likely to help classify species correctly?

#### Preliminary Trend Inspection

```{r, results = TRUE}
#reading in the data
palmetto <- read_csv(here("data", "palmetto.csv")) %>%
  #selecting species and chosen variables
  select(species, height:green_lvs) %>% 
  #making species a factor
   mutate(species = as.factor(species)) %>% 
  #dropping na and other 
  drop_na()






#making a plot with height and canpy width 
plot1 <- ggplot(data = palmetto, aes(x = height, y = width)) +
  geom_point(aes(color = species)) +
  ggtitle("") + 
 labs(x = "Plant Height (cm)", y = "Canopy Width (cm)") +
  scale_color_discrete(name = "Species", labels = c("1" = " Serenoa repens ", "2" = "Sabal etonia"))

#making a plot with height and canopy length
plot2 <- ggplot(data = palmetto, aes(x = height, y = length)) +
  geom_point(aes(color = species))  +
  ggtitle("") + 
 labs(x = "Plant Height (cm)", y = "Canopy Length (cm)") +
  scale_color_discrete(name = "Species", labels = c("1" = " Serenoa repens ", "2" = "Sabal etonia"))


#making a plot with height and leaf count
plot3 <- ggplot(data = palmetto, aes(x = height, y = green_lvs)) +
  geom_point(aes(color = species)) +
  ggtitle("") + 
 labs(x = "Plant Height (cm)", y = "Green Leaf Count") + 
    scale_color_discrete(name = "Species", labels = c("1" = " Serenoa repens ", "2" = "Sabal etonia"))



#patching together the plots
plot_patch <- plot3  + plot2  / plot1 +plot_layout(guides = 'collect') & theme( legend.position = 'bottom') 


#adding plot annotations A and B
 plot_patch +  plot_annotation("Preliminary Inspection of Predictor Variables",
  tag_levels = "A") 
```

**Figure 1**: Preliminary data observations across variables.

**A:** Comparing plant height and total green leaf count. *S. repens* appears to higher green leaf counts compared to *S. etonia* per unit of height.

**B:** Comparing plant height and canopy length. *S. repens* appears to have slightly smaller canopy length measurements compared to *S. etonia* per unit of height.

**C:** Comparing plant height canopy width. Appears to be much overlap between species.

Based off of the initial data visualization, canopy length and green leaf count are going to be the two most predicting variables determining the type of species. Height and width show significant overlap across both species where as canopy length and green leaf count yield two distinct patterns for both species.

#### Binary Logistic Regression Modeling

c.  A section in which you perform binary logistic regression to determine the probability of a plant being either Serenoa repens or Sabal etonia based on several predictor variables. Perform the analysis twice, using cross validation to compare two models:

    ```{r}
    #creaitng the model input for model 1
    f1 <- species ~ height + length + width + green_lvs


    plant_blr1 <- glm(formula = f1,
                    data = palmetto,
                    family = "binomial")

    #visualizing the model
    blr1_tidy <- broom::tidy(plant_blr1)
    blr1_tidy





    #creaitng the model input for model 2
    f2 <- species ~ height + width + green_lvs


    plant_blr2 <- glm(formula = f2,
                    data = palmetto,
                    family = "binomial")

    #visualizing the model
    blr2_tidy <- broom::tidy(plant_blr2)
    blr2_tidy




    ```


#### need to work on this. has to got to the end of task 2 for part e
```{r}
# converting the odds to the probability of being 1 or 2   
blr1_fitted <- plant_blr1 %>%
  broom::augment(type.predict = "response")   %>% 
  mutate(species_pred = ifelse(.fitted > 0.5, "2" , "1"))

blr2_fitted <- plant_blr2 %>%
  broom::augment(type.predict = "response")

```

#### Grphing the Models

```{r}
ggplot_blr1 <- ggplot(data = blr1_fitted, aes(x = green_lvs, y = .fitted)) +
  # add aes(shape = species) to compare probability with actual
  geom_point(aes(color = species)) +
  # add geom_smooth to show general fit
  geom_smooth() +
  labs(x = "Green Leaves Count",
   	   y = "Probability of outcome S. Etonia") + 
    scale_color_discrete(name = "Species", labels = c("1" = " Serenoa repens ", "2" = "Sabal etonia"))


ggplot_blr2 <- ggplot(data = blr1_fitted, aes(x = width, y = .fitted)) +
  # add aes(shape = species) to compare probability with actual
  geom_point(aes(color = species)) +
  # add geom_smooth to show general fit
  geom_smooth() +
  labs(x = "Canopy Width (cm)",
   	   y = "Probability of outcome S. etonia") + 
    scale_color_discrete(name = "Species", labels = c("1" = " Serenoa repens ", "2" = "Sabal etonia"))

#patching together the plots
blr_patch <- ggplot_blr1  + ggplot_blr2  +plot_layout(guides = 'collect') & theme( legend.position = 'bottom') 

#adding plot annotations A and B
blr_patch +  plot_annotation("Binary Least Regression Models",
  tag_levels = "A") 

```

### AIC and BIC

```{r, results=TRUE}
# Listing the AIC values
AIC_list<-c(AIC(plant_blr1), AIC(plant_blr2))
BIC_list <- c(BIC(plant_blr1), BIC(plant_blr2))

# merging data and selecting specific columns
model_output <-rbind(data.frame(glance(plant_blr1)),data.frame(glance(plant_blr2))) 

# calculate delta AIC or the distance in AIC from the lowest value (the best model) and the other models.
model_output <- mutate(model_output, delta.AIC = AIC-min(AIC_list))
model_output <- mutate(model_output, delta.BIC = BIC-min(BIC_list))
model_output$model<-c( "Model 1","Model 2")
model_output<-model_output[,c("model", "AIC", "delta.AIC", "BIC", "delta.BIC" )]


kable(model_output, format = "markdown", digits = 3, caption = "*Figure 2*: Model outputs showing the AIC/BIC values and delta AIC/BIC values for both models.") 
 
```

#### 10 fold

#### K fold

```{r}
set.seed(123)


n_folds <- 10
fold_vec <- rep(1:n_folds, length.out = nrow(palmetto))
palmetto_kfold <- palmetto %>%
  mutate(fold = sample(fold_vec, size = n(), replace = FALSE))




results_df <- data.frame()
pred_acc <- function(x, y) {
  accurate <- ifelse(x == y, 1, 0)
  return(mean(accurate, na.rm = TRUE))
}

for(i in 1:n_folds) {
  kfold_test <- palmetto_kfold %>%
    filter(fold == i)
  kfold_train <- palmetto_kfold %>%
    filter(fold != i)
  
  kfold_blr1 <- glm(f1, data = kfold_train, family = 'binomial')
  kfold_blr2 <- glm(f2, data = kfold_train, family = 'binomial')
  kfold_pred <- kfold_test %>%
    mutate(blr1 = predict(kfold_blr1, kfold_test, type = 'response'),
           blr2 = predict(kfold_blr2, ., type = 'response')) %>%
    mutate(pred1 = ifelse(blr1 > 0.50, '2', '1'),
           pred2 = ifelse(blr2 > 0.50, '2', '1'))
  kfold_accuracy <- kfold_pred %>%
    summarize(blr1_acc = pred_acc(species, pred1),
              blr2_acc = pred_acc(species, pred2))
  
  results_df <- bind_rows(results_df, kfold_accuracy)
}


results_df %>%
  summarize(blr1_acc = mean(blr1_acc),
            blr2_acc = mean(blr2_acc))

kable(results_df,  format = "markdown", digits = 3,caption = "*Figure 3*:")
```

#### train entire data set on it

Our final model: `r equatiomatic::extract_eq(plant_blr1, wrap = TRUE)`

and with coefficients in place: `r equatiomatic::extract_eq(plant_blr1, wrap = TRUE, use_coefs = TRUE)`




#run glm for final model








e.  A section that evaluates how successfully this model would "classify" a plant as the correct species, using a 50% cutoff (e.g. if the probability is \>=50% that it is species A, then it would be classified as species A). Use broom::augment() to find the probabilities (instead of log-odds) for each plant in the original dataset, then add a column for which species your model would classify that plant as (using a 50% cutoff) based on the included predictor variables. The outcome should be a finalized table showing, for each species, how many plants in the original dataset would be correctly classified and how many were incorrectly classified by the model, as well as an additional column with "% correctly classified". Add a table caption above the table, and a 1-2 sentence conclusion paragraph after.



blr_f1_pred <- adelie_chinstrap %>%
  mutate(predict(blr_tidyfit_f1, .),
         predict(blr_tidyfit_f1, ., type = 'prob'))

blr_f1_pred %>%
  roc_curve(truth = species, .pred_Adelie) %>%
  autoplot()

blr_f1_pred %>%
  roc_auc(truth = species, .pred_Adelie)

### Students repeat for blr_tidyfit_f2 and compare!
blr_f2_pred <- adelie_chinstrap %>%
  mutate(predict(blr_tidyfit_f2, .),
         predict(blr_tidyfit_f2, ., type = 'prob'))

blr_f2_pred %>%
  roc_curve(truth = species, .pred_Adelie) %>%
  autoplot()

blr_f2_pred %>%
  roc_auc(truth = species, .pred_Adelie)


Final table only needed for final model 

#based on the probablity colomn (fitted)
use this to create match species 1 =species 1 --> True
species 1 - species 2 --> False
make another column saying the proportion of True to False 



