---
title: "Task 2"
author: "Ray Hunter"
date: "2023-02-12"
output:   
  html_document:
    code_folding: hide
---

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo=TRUE, warning=FALSE, message=FALSE,  results=FALSE)
library(purrr)
library(tidyverse)
library(janitor)
library(Metrics)
library(cowplot)
library(here)
library(kableExtra)
library(Metrics)
```

## Task 2

#### Introduction:

The purpose this task is to create a model that accurately fits lizard phenotype traits to a data set of lizard observations. To do this, we use a non linear least squares model to estimate the parameters of a snout length to weight model. All data was collected in New Mexico from 1989-2006 across twenty different species (Ligthfoot and Whitford, 2020).  

**Dats Source:** Lightfoot, D. and W.G. Whitford. 2020. Lizard pitfall trap data from 11 NPP study locations at the Jornada Basin LTER site, 1989-2006 ver 37. Environmental Data Initiative. https://doi.org/10.6073/pasta/4a6e258fb49c31e222ecbbcfd128967f



#### creating an NLS model 
```{r}
#reading in data and wrangling
mdl1_data<-  read_csv(here("data", "lizards.csv")) %>% 
  clean_names() %>% 
  #taking th log of values 
  mutate(log_weight = log(weight), log_length = log(sv_length)) 
  

mdl1_data$sex[mdl1_data$sex=="F"] <- "female"
mdl1_data$sex[mdl1_data$sex=="M"] <- "male"
 
```

**1:** Model equation for lizard weigth ($W$) and snout length ($SVL$) where $a$ and $b$ are parameters that need to be fitted.

$$  W =a(SVL)^b $$


**2:** Because the data has an exponential nature, we need to use a log transformation to create a linear model using ordinary least squares (OLS). This then gives us the intercepts and slope coefficient values of the model that can be applied to the original equation to create a non linear least squares (NLS) model.*

$$ log(W)= log(a) + b*log(SVL)$$
```{r }
#running an OLS model
guess_mod1 <- lm(log_weight ~log_length, data = mdl1_data) 


#creating a funciton that calculates the original equation
func1<-function(a,b,length){
 out= a*(length^b)
return(out)
}




#creating and NLS model 
nls_mdl1=nls(weight ~func1(a,b, sv_length),
                  data=mdl1_data,
                  start=list(a=exp(-8.25), b=2.479417),
                  trace=TRUE)
summary(nls_mdl1)
```

**3:** We then extract the coefficients from the OLS model and apply them to a function that can be used to create a non-linear least regressions NLS model using the original equation.

$$ W = e^{-8.25} (SVL)^{2.48}$$



```{r, fig.cap="Figure 1: Non linear least squars (NLS) model showing the relation ship between snout-vent length (mm) and predicted body weigth (g) fitted to actual body weight measurements for all lizards. "}
#creating a new data set with model predicitions
mdl1_predict<-mdl1_data %>% 
  mutate(predict=predict(nls_mdl1,newdata=.))

#visualizing the NLS
ggplot1 <- ggplot(data=mdl1_predict)+
  geom_point(aes(x=sv_length,y=weight,colour = sex)) + 
  ggtitle("NLS Data Predictions") + 
 labs(x = "Snout-Vent Length (mm)", y = "Body Weight (g)")+
  
    geom_smooth(aes(x=sv_length,y=predict), color='black') +
   facet_wrap(~sex)
ggplot1
```


```{r, results=TRUE}
#### NLS model output

broom::tidy(nls_mdl1) %>% 
  kable(caption = "Figure 2: NLS model showing the parameter estimates for *a* an *b*. These parametere are used in the model to predict body mass (g) from snot-vent length (mm).") %>% 
  kable_material()
```






































#### second NLS


```{r}
#wrangle new data set
mdl2_data <- mdl1_data %>% 
  filter(spp =="CNTI", sex=="male")
```


#### NLS Model Creation and Figure
```{r}

#finding parameter values from lm of log values
guess_mod2 <- lm(log_weight ~log_length, data = mdl2_data) 
summary(guess_mod2)


#creating NLS with parameters found from log lm 
nls_mdl2=nls(weight ~func1(a,b, sv_length),
                  data=mdl2_data,
                  start=list(a=exp(-9.0240), b=2.6981),
                  trace=TRUE)


#creating a predict df for model 2 
mdl2_predict<-mdl2_data %>% 
  mutate(predict=predict(nls_mdl2,newdata=.)) 
```

```{r, fig.cap="Figure 3: Non linear least squars (NLS) model showing the relation ship between snout-vent length (mm) and predicted body weigth (g) fitted to actual body weight measurements. NLS model 1 (root mean squared error = 1.562) fitting parameters to all lizards is shown in red and NLS model 2 (root mean squarred error = 3.349) fitting parameters for just male Western Whiptail lizards (*Cnemidophorus tigrisatus*) is in blue. Becasue the NLS 1 model has a lower RMSE, it predicts the data better and therefore should be used over the NLS model 2 to fit the *C. tigrisatus* data." }





ggplot2 <- ggplot(data=mdl2_predict)+ xlim(0,100) +ylim(0,40) +
  geom_point(aes(x=sv_length,y=weight)) + 
  ggtitle("NLS Data Predictions") + 
 labs(x = "Snout-Vent Length (mm)", y = "Body Weight (g)")+
    geom_smooth(show.legend=TRUE, aes(x=sv_length,y=predict), color='blue') +
  geom_smooth(data=mdl1_predict, show.legend=TRUE, aes(x=sv_length, y=predict), color="red") +
  scale_colour_manual(name="legend", values=c("blue", "red")) +
  scale_linetype_discrete(labels=c("NLS 1", "NLS 2"))
  
ggplot2






#calculating rmse for two models
rmse_mdl1 <- rmse(mdl1_predict$weight, mdl1_predict$predict)
rmse_mdl2 <- rmse(mdl2_predict$weight, mdl2_predict$predict)


```























Do I need to make a model #3 using the model 1 (total data set) a and b parameters and apply it to the model 2 data set (just whip tail)? if so... but not workign becasue model 2 and model 3 NLS are the same
```{r}
#creating a model 3 nl whihc is model 1 nls using data from mdl2
nls_mdl3=nls(weight ~func1(a,b, sv_length),
                  data=mdl2_data,
                  start=list(a=exp(-8.25), b=2.479417),
                  trace=TRUE)

#creating a predict df for model 3 using model 2 (just male Western Whiptail data)
mdl3_predict<-mdl2_data %>% 
  mutate(predict=predict(nls_mdl3,newdata=mdl2_data)) 
```





























































